Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-10 paper/main.tex
Updates

In Algorithm~\ref{algo:expert}, in each iteration we add a new edge to the DAG
chosen by Algorithm~\ref{algo:expand}. This edge is chosen at random depending
on the order of iteration in Algoirthm~\ref{algo:expand}. A more optimal way to
add these edges would be to choose the pairs of variables that could help
improve the fit the most in each step. One potential greedy approach to do this
would be to select the variable pair $ X $ and $ Y $ that have the highest
amount of observed conditional correlation in the data. In
Algorithm~\ref{algo:ranked_expand}, we slightly modify
Algorithm~\ref{algo:expand} to show this approach. In this case we first go
through all potential edges that can be added to the DAG in each step, and rank
them based on the conditional association between them given the current DAG.
This gives us a measure of how much of the observed correlation between them is
not explained by the current model. We then choose the pair of variables the
highest conditional association. This modification allows us to prioritize pair
of variables to quickly improve the fit of the model. For the conditional
measure of association, we use a measures outlined in
Section~\ref{sec:background} depending on the data type of the variables.

>>>>

The \textsc{ExpertInLoop} algorithm adds a new edge in each iteration. The edge
to be added is selected using the \textsc{Expand} algorithm, which returns a
potential edge at random based on the iteration order in the algorithm.
However, if the residual association between the selected variables is low,
adding the edge may result in only a marginal improvement in the model's
overall fit. As the algorithm requires manual intervention in each iteration in
terms of domain knowledge, it would be beneficial to prioritize the edges that
improve the model's fit the most.

Therefore, an alternative to selecting an edge at random is to rank potential
edges based on the residual association between the variables given the DAG.
For quantifying the residual association between variables, we can use the
effect size measures of the CI test. So, for a d-separation query $ D_G(X, Y,
\mathbf{Z}) $, we can quantify the residual association between $ X $ and $ Y $
conditioned on $ \mathbf{Z} $ as the effect size of the CI test $ X \ci Y |
\mathbf{Z} $. A greedy approach could be to choose the edge that has the
highest residual association between the variables.
Algorithm~\ref{algo:ranked_expand} outlines this approach. This approach has
the benefit that it prioritizes edges that improve the model fit the most.
